---
title: "script_for_AJMG"
author: "DVM Bishop"
date: "07/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# many of these functions copied from CNV_for_protocol
require(tidyverse)
require(stringr)
require(plyr)
require(yarrr)
require(knitr)
library(Hmisc) #for correlation matrix
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(scales)
require(sandwich)
require(msm)
require(lemon) #for glegend
require(MASS)
require(simglm)
require(conflicted) #flags up conflicted functions
require(sandwich)
```

## Read and merge files

Github repository is CNV_Hayley

R project is CNV_Hayley.Rproj on deevybee_repo

Protocol is CNV_for_protocol.rmd from Jun 16 2018

Manuscript draft is with oscci manuscripts in folder hayley cnv, as Trisomy_CNV_AJMG_v1_DB 

Data on CNVs now in same folder as R project; original xls file from Dianne, with parental concern variable, is saved as 2 worksheets in csv – one for autosomes and one for X chromosome
all_samples_results_parco.csv
all_samples_results_Xchrom.csv

These need to have name codes changed to match our codes by stripping off DB and making other minor modifications.

Remember to set working directory to source file location

NB famsplangconc coded as 1 if any concern about speech/language in either twin. (May be co-twin to the proband) - some of these were coded N/A in Di's file - have updated these manually.


```{r readfiles}
cnvdat<-read.csv('all_samples_results_parco_corrected.csv',stringsAsFactors=FALSE)
cnvdatx<-read.csv('all_samples_results_Xchrom.csv',stringsAsFactors=FALSE)

identical(cnvdat$id,cnvdatx$id) #OK - original files were not identical- found mismatch

genderdat<-read.csv('TwinsData_gender.csv')
phenodat<-read.csv('unblinded_pheno_sct_twin.csv',stringsAsFactors=FALSE)
#This file was created using 'geno_pheno_analysis_190215.Rmd' - i.e. main script for Newbury et al paper. In that script, IDs are blinded, but here we need to match up with CNV data, so retain the codes.

#need to strip out correct ID from the cnv files
#first of all remove _rpt from end of name
#also remove DB from start
#using cnvdatx as this seems more accurate, but ultimately should agree with cnvdat



for (i in 1:nrow(cnvdatx)){
  cnvdatx$id[i] <-  str_replace(cnvdatx$id[i],"_rpt", "")
  cnvdatx$id[i] <-  str_replace(cnvdatx$id[i],"DB", "")
    cnvdat$id[i] <-  str_replace(cnvdat$id[i],"_rpt", "")
  cnvdat$id[i] <-  str_replace(cnvdat$id[i],"DB", "")
}

#for SCTs, just retain first 3 digits
myrows<-which(cnvdatx$Group != 'twin')

cnvdatx$id[myrows]<-str_sub(cnvdatx$id[myrows],1,3)
cnvdat$id[myrows]<-str_sub(cnvdatx$id[myrows],1,3)
#NB these ids are characters, so need to alter phenodat,where id is numeric
#Need to add leading zeroes, so, e.g. 1 becomes 001

phenodat$id<-as.character(phenodat$record_id)
w<-which(as.numeric(phenodat$record_id)<10)
phenodat$id[w]<-paste0('0',phenodat$id[w])
w<-which(as.numeric(phenodat$record_id)<100)
phenodat$id[w]<-paste0('0',phenodat$id[w])

#order all files by id
genderdat <- genderdat[order(genderdat$record_id),] 
phenodat <- phenodat[order(phenodat$id),] 
cnvdat <- cnvdat[order(cnvdat$id),] 
cnvdatx <- cnvdatx[order(cnvdatx$id),] 
#align files and bolt together

#Just check we do have same cases in same rows
for(i in 1:nrow(cnvdat)){
  if(cnvdat$id[i] != cnvdatx$id[i]){
    print(cnvdat$id[i])
  }
  }



w<-which(phenodat$id %in% cnvdatx$id)

cnvpheno<-cbind(phenodat[w,2:44],cnvdat[,2:11],cnvdatx[,2:8])

cnvpheno$female<-0
w4<-which(cnvpheno$trisomy==1)
cnvpheno$female[w4]<-1

w3<-which(cnvpheno$Group=='twin')
w2<-which(genderdat$record_id %in% cnvpheno$id) #twins only!

cnvpheno$female[w3]<-genderdat$female[w2] #twins who are in cnvpheno assigned gender from genderdat

cnvpheno$groupfem<-paste0(cnvpheno$Group,cnvpheno$female)

#Do some renaming of cols
for (i in 54:60){
  colnames(cnvpheno)[i]<-paste0(colnames(cnvpheno)[i],'X')
}
cnvpheno$is.trisomy <-1
w<-which(cnvpheno$Group=='twin')
cnvpheno$is.trisomy[w]<-0 


```

We now have a file with the phenotype and genotype data together.  
First thing that is needed is Table 1, showing Ns in relation to ascertainment bias.
```{r table1}

cnvpheno$bias<-cnvpheno$famsplangconc #default from twin code for family bias

w<-which(cnvpheno$Group !='twin')
cnvpheno$bias[w]<-0 #default no bias for trisomies
w<-c(which(cnvpheno$why_tested==2),which(cnvpheno$why_tested==3))
cnvpheno$bias[w]<-1

mytab<- table(cnvpheno$groupfem,cnvpheno$bias)

# Add a variable to identify those with pathogenic cnvs
patholist <-c('085','613A','244','679A','066','339','348','257','134','355','668B','869B','664B','874A','131','639A')
w<-which(cnvpheno$id %in% patholist)
cnvpheno$pathoCNV <-0
cnvpheno$pathoCNV[w] <-1

write.csv(cnvpheno,'cnvpheno.csv',row.names=FALSE) #saved copy of file

```
The derivation of the global neurodevelopmental index is now provided in the paper in Table 2.
(In this version we use original scores, so high score represents poor functioning. 



```{r cnv.check, echo=TRUE}
##############################################################################


plot(jitter(cnvpheno$total_cnvs),cnvpheno$sum_pLi)
abline(a=0,b=1) # not sure how sum_pLi can be bigger than total_cnvs. But there's a different variable called no_pLi. Need to check with Hayley/Di

```

```{r original.cnvplot}
# Create a censored version of sum_pLi to fix top of scale at 7
# Censored variable is sum_pLi_C
#Function borrowed and adapted from the 'scales' package. This applies the censoring to the two potentially interesting variables.
#str(cnvpheno) #uncomment to inspect variables
censor_PT<-function (x, limit = 7, only.finite = TRUE) 
{
  force(range)
  finite <- if (only.finite) 
    is.finite(x)
  else TRUE
  x[finite & x > limit] <- limit+1
  x
}

cnvpheno$sum_pLi_C<-censor_PT(cnvpheno$sum_pLi,limit=7,only.finite = F)
#cnvpheno$total_Kbp_C<-censor_PT(cnvpheno$total_Kbp,limit=3000,only.finite = F)
#mytab<-table(cnvpheno$sum_pLi,cnvpheno$sum_pLi_C)
#plot(cnvpheno$sum_pLi,cnvpheno$sum_pLi_C) #show effect of censoring

#hist(cnvpheno$sum_pLi_C) #NB still v skewed



#-----------------------------------------------------------------------------#
p1 <- ggplot(cnvpheno,aes(x=sum_pLi,y=sum_pLi_C)) + geom_point() +
  scale_x_continuous(expand=c(0.02,0)) +
  scale_y_continuous(expand=c(0.02,0)) +
  theme_bw() +
  theme(legend.position="none",plot.margin=unit(c(1,1,1,1),"points"))

theme0 <- function(...) theme( legend.position = "none",
                               panel.background = element_blank(),
                               panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(),
                               panel.margin = unit(0,"null"),
                               axis.ticks = element_blank(),
                               axis.text.x = element_blank(),
                               axis.text.y = element_blank(),
                               axis.title.x = element_blank(),
                               axis.title.y = element_blank(),
                               axis.ticks.length = unit(0,"null"),
                               axis.ticks.margin = unit(0,"null"),
                               panel.border=element_rect(color=NA),...)

p2 <- ggplot(cnvpheno,aes(x=sum_pLi)) + 
  geom_density(alpha=0.5) + 
  scale_x_continuous(breaks=NULL,expand=c(0.02,0)) +
  scale_y_continuous(breaks=NULL,expand=c(0.00,0)) +
  theme_bw() +
  theme0(plot.margin = unit(c(0,-1,0.5,1.5),"lines")) 

p3 <- ggplot(cnvpheno,aes(x=sum_pLi_C)) + 
  geom_density(alpha=0.5) + 
  coord_flip()  + 
  scale_x_continuous(labels = NULL,breaks=NULL,expand=c(0.02,0)) +
  scale_y_continuous(labels = NULL,breaks=NULL,expand=c(0.00,0)) +
  theme_bw() +
  theme0(plot.margin = unit(c(0,0,1.75,1),"lines"))

#mylegend<-g_legend(p1)  #This throws error, even after installing lemon

grid.arrange(arrangeGrob(p2,ncol=2,widths=c(3,1)),
             arrangeGrob(p1,p3,ncol=2,widths=c(3,1)),
             heights=c(1,3))


cnvpheno$is.trisomy<-as.factor(cnvpheno$is.trisomy)
#-----------------------------------------------------------------------------#
```


```{r visualise.dists, echo=FALSE}
#-----------------------------------------------------------------------------#

# Scatter plot with marginal histograms 
# NB this needs some modification
# I have used the raw rather than censored values for the plot (but they only go up to 8?)
# Have taken log for kBp to aid visualisation
# Added legend to bottom, but now the scales for the density plots are messed up...
# Used colour to depict twin/SCT and shape to depict bias and whether patho CNV

# Would like to include jitter on x-axis, but my attempt to do this just duplicated points!

# Create factor variables for group and bias
cnvpheno$groupfac<-factor(cnvpheno$is.trisomy)

cnvpheno$bias_path<-as.factor(10*cnvpheno$bias+cnvpheno$pathoCNV)
shapelist<-c(1,19,0,15) #shapes are unfilled circle, filled circle, unfilled sq, filled sq
#This variable allows mapping of shape and fill as follows:
# No bias, no pathoCNV = unfilled circle
# No bias, pathoCNV = filled circle
# Bias, no pathoCNV = unfilled square
# Bias, pathoCNV = filled square


p1 <- ggplot(cnvpheno,aes(x=sum_pLi,y=log(total_Kbp),colour=is.trisomy,shape=bias_path))+
  scale_shape_manual(values=shapelist)+
   scale_color_manual(values=c('gray21','red'))+
  geom_point()+
  scale_x_continuous(expand=c(0.02,0)) +
  scale_y_continuous(expand=c(0.02,0)) +

  theme_bw() +
  theme(legend.position="bottom",plot.margin=unit(c(1,1,1,1),"points"))

theme0 <- function(...) theme( legend.position = "none",
                               panel.background = element_blank(),
                               panel.grid.major = element_blank(),
                               panel.grid.minor = element_blank(),
                               panel.margin = unit(0,"null"),
                               axis.ticks = element_blank(),
                               axis.text.x = element_blank(),
                               axis.text.y = element_blank(),
                               axis.title.x = element_blank(),
                               axis.title.y = element_blank(),
                               axis.ticks.length = unit(0,"null"),
                               axis.ticks.margin = unit(0,"null"),
                               panel.border=element_rect(color=NA),...)

p2 <- ggplot(cnvpheno,aes(x=sum_pLi,colour=factor(is.trisomy),fill=factor(is.trisomy))) + 
  geom_density(alpha=0.5) + 
  scale_color_manual(values=c('gray21','red'))+
  scale_fill_manual(values=c('gray21','red'))+
  scale_x_continuous(breaks=NULL,expand=c(0.02,0)) +
  scale_y_continuous(breaks=NULL,expand=c(0.00,0)) +
  theme_bw() +
  theme0(plot.margin = unit(c(0,0,0.5,3),"lines")) 

p3 <- ggplot(cnvpheno,aes(x=log(total_Kbp),colour=factor(is.trisomy),fill=factor(is.trisomy))) + 
  geom_density(alpha=0.5) + 
  scale_color_manual(values=c('gray21','red'))+
   scale_fill_manual(values=c('gray21','red'))+
  coord_flip()  + 
  scale_x_continuous(labels = NULL,breaks=NULL,expand=c(0.02,0)) +
  scale_y_continuous(labels = NULL,breaks=NULL,expand=c(0.00,0)) +
  theme_bw() +
  theme0(plot.margin = unit(c(0,0,1.75,1),"lines"))



grid.arrange(arrangeGrob(p2,ncol=2,widths=c(3,1)),
             arrangeGrob(p1,p3,ncol=2,widths=c(3,1)),
             heights=c(1,3))


##############################################################################
```


```{r corrcnv,echo=TRUE}
mycor <- rcorr(as.matrix(cbind(cnvpheno$total_cnvs,cnvpheno$total_Kbp,cnvpheno$sum_pLi_C)),type='pearson')
rtab<-round(mycor$r,3)
colnames(rtab) <- c('total cnv','total Kbp','sum pLi')
rownames(rtab) <- c('total cnv','total Kbp','sum pLi')

knitr::kable(rtab,Caption='Correlations between CNV measures')
```

Next we consider how well the different measures differentiate between the groups. Because the data is non-normal, a Wilcoxon rank sum test is used. In a preliminary analysis, we analyse each dependent measure separately.

```{r comparegp,echo=FALSE}
##############################################################################

#Multiple testing: Wilcoxon sum rank test, one-tailed
dowilcox <- function(mydf){
BF_correct_alpha<-0.05/3
#print('Total CNVs')
w1<-wilcox.test(total_cnvs~is.trisomy,data=mydf,alternative=c("less"),conf.int=TRUE)
#print(list(p.value=w1$p.value,corrected.alpha=BF_correct_alpha,res=w1$p.value<=BF_correct_alpha))
#print('Total Kbp')
w2<-wilcox.test(total_Kbp~is.trisomy,data=mydf,alternative=c("less"),conf.int=TRUE)
#print(list(p.value=w2$p.value,corrected.alpha=BF_correct_alpha,res=w2$p.value<=BF_correct_alpha))
#print('sum pLi')
w3<-wilcox.test(sum_pLi_C~is.trisomy,data=mydf,alternative=c("less"),conf.int=TRUE)
#print(list(p.value=w3$p.value,corrected.alpha=BF_correct_alpha,res=w3$p.value<=BF_correct_alpha))

#Corrections for multiple testing

# Given correlations between the CNV measures, the Benjamini-Yekutieli (BY) correction is applied. We see that the pattern of significance among the tests remains unchanged.

##
pvals<-round(c(w1$p.value,w2$p.value,w3$p.value),2)
BY = p.adjust(pvals, "BY")
res = cbind(pvals, BY=round(BY, 3))
colnames(res)<-c("Uncorrected","Benjamini-Y")
rownames(res)<-c('Total CNVs','Total Kbp',"sum pLi")
return(res)}

```

```{r makeresultstabs}

# Divide data into 2 sets; with and without bias
w<-which(cnvpheno$bias==1)
biasset<-cnvpheno[w,]
nobiasset<-cnvpheno[-w,]

res <-dowilcox(cnvpheno)
knitr::kable(res, caption='All cases')

res1 <-dowilcox(nobiasset)
knitr::kable(res1, caption='Low bias cases')

by_group<-cnvpheno %>% group_by(is.trisomy)
groupmeans<-dplyr::summarise(by_group,
  mean_cnvs = round(mean(total_cnvs),2),
   sd_cnvs = round(sd(total_cnvs),2),
    mean_Kbp = round(mean(total_Kbp),1),
   sd_Kbp = round(sd(total_Kbp),1),
  mean_pLi = round(mean(sum_pLi),3),
  sd_pLi = round(sd(sum_pLi),3)
)
groupmeans

```




## Phenotype analysis
We have three possible phenotypes that were used by Newbury et al (2018). We decided *a priori* that the measure of global neurodevelopmental impairment is optimal, because previous research on CNVs implicates more severe conditions such as ASD and intellectual disability, which are captured by this measure.




## Poisson regression
Poisson regression is the optimal method, as the dependent variable is a count. This will naturally allow for the lower bound of zero.

Reference for analysis: https://stats.idre.ucla.edu/r/dae/poisson-regression/
```{r poisson, echo=TRUE}
#-----------------------------------------------------------------------------#
summary(m1 <- glm(global_neurodev ~ sum_pLi + is.trisomy + sum_pLi:is.trisomy, family="poisson", data=cnvpheno))

#check the assumption that mean and variance ar roughly equal. If not, we can use robust estimates. I think this will make little difference in our case but included for completeness.

cov.m1 <- vcovHC(m1, type="HC0")
std.err <- sqrt(diag(cov.m1))
r.est <- cbind(Estimate= coef(m1), "Robust SE" = std.err,
               "Pr(>|z|)" = 2 * pnorm(abs(coef(m1)/std.err), lower.tail=FALSE),
               LL = coef(m1) - 1.96 * std.err,
               UL = coef(m1) + 1.96 * std.err)

r.est

#test the effect of dropping is.trisomy. We anticipate that this will be impair model fit.
m2 <- update(m1, . ~ . - is.trisomy)
## test model differences with chi square test
anova(m2, m1, test="Chisq")

#-----------------------------------------------------------------------------#
```
Added bits by DB, Nov 2019.
Just recreating Hayley's analysis of burden.

```{r burdentable}


#make data frame to hold summary results
mydf<-data.frame(matrix(NA,nrow=10,ncol=7))
colnames(mydf)<-c('Variable','All SCT','All twin','Low bias SCT','Low bias twin','High bias SCT','High bias twin')

temptab<-table(cnvpheno$is.trisomy,cnvpheno$bias)
mydf[1,1]<-'N'
mydf[1,4]<-temptab[2,1]
mydf[1,5]<-temptab[1,1]
mydf[1,6]<-temptab[2,2]
mydf[1,7]<-temptab[1,2]
mydf[1,2]<-mydf[1,4]+mydf[1,6]
mydf[1,3]<-mydf[1,5]+mydf[1,7]

myvarlist<-c('total_cnvsX','total_cnvs','total_Kbp')
varcols<-which(colnames(cnvpheno)%in% myvarlist)
 myagg<-aggregate(cnvpheno[,varcols],by=list(cnvpheno$is.trisomy,cnvpheno$bias),FUN=sum,na.rm=TRUE)
 
 colnames(myagg)[1:2]<-c('Trisomy','Bias')
 myagg$Trisomy<-as.character(myagg$Trisomy)
 myagg[5,1]<-'All twin'
myagg[5,2]<-'-'
 myagg[6,1]<-'All trisomy'
myagg[6,2]<-'-'
 myagg[5,3]<-sum(myagg[c(1,3),3])
  myagg[5,4]<-sum(myagg[c(1,3),4])
   myagg[5,5]<-sum(myagg[c(1,3),5])
    myagg[6,3]<-sum(myagg[c(2,4),3])
  myagg[6,4]<-sum(myagg[c(2,4),4])
   myagg[6,5]<-sum(myagg[c(2,4),5])
 
   myagg$N[1:6] <- as.numeric(mydf[1,c(5,4,7,6,3,2)])
    myagg$avgburden<- myagg$total_cnvs/myagg$N
      myagg$avgspan<- myagg$total_Kbp/myagg$N
    
   
 
maketable<-function(mydf,mylabel,myrow,myvar1){
thiscol<-which(colnames(cnvpheno)==myvar1)
  mydf[myrow,1]<-mylabel

  mydf[myrow,2]<-sum(dplyr::filter(cnvpheno[,thiscol],myvar1==1))
  mydf[myrow,3]<-sum(dplyr::filter(cnvpheno,myvar1==0))
    mydf[myrow,4]<-nrow(dplyr::filter(nobiasset,myvar1==1))
  mydf[myrow,5]<-nrow(dplyr::filter(nobiasset,myvar1==0))
    mydf[myrow,6]<-nrow(dplyr::filter(biasset,myvar1==1))
  mydf[myrow,7]<-nrow(dplyr::filter(biasset,myvar1==0))
 
}

mylabel='N'
myrow=1
myvar1='total_cnvs'
maketable(mydf,mylabel,myrow,myvar1)
```


```{r bit}
#do pirate-plots
pirateplot(log(1+cnvpheno$total_cnvs)~cnvpheno$bias*cnvpheno$is.trisomy,data=cnvpheno)
pirateplot(log(1+cnvpheno$total_Kbp)~cnvpheno$bias*cnvpheno$is.trisomy,data=cnvpheno)

#As explained in the paper, the X chromosome is a bit of a nightmare if you have 3 Xs!
pirateplot(log(1+log(1+cnvpheno$total_KbpX))~cnvpheno$bias*cnvpheno$is.trisomy*cnvpheno$female,data=cnvpheno)
```

```{r ind.cnvs}
mycnvs<-read.csv()

```
