---
title: "script_for_AJMG"
author: "DVM Bishop"
date: "07/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(stringr)
require(plyr)
require(yarrr)
```

## Read and merge files

Github repository is CNV_Hayley

R project is CNV_Hayley.Rproj on deevybee_repo

Protocol is CNV_for_protocol.rmd from Jun 16 2018

Manuscript draft is with oscci manuscripts in folder hayley cnv, as Trisomy_CNV_AJMG_v1_DB 

Data on CNVs now in same folder as R project; original xls file from Dianne, with parental concern variable, is saved as 2 worksheets in csv – one for autosomes and one for X chromosome
all_samples_results_parco.csv
all_samples_results_Xchrom.csv

These need to have name codes changed to match our codes by stripping off DB and making other minor modifications.

Remember to set working directory to source file location

NB famsplangconc coded as 1 if any concern about speech/language in either twin. (May be co-twin to the proband) - some of these were coded N/A in Di's file - have updated these manually.


```{r readfiles}
cnvdat<-read.csv('all_samples_results_parco_corrected.csv',stringsAsFactors=FALSE)
cnvdatx<-read.csv('all_samples_results_Xchrom.csv',stringsAsFactors=FALSE)

identical(cnvdat$id,cnvdatx$id) #OK - original files were not identical- found mismatch

genderdat<-read.csv('TwinsData_gender.csv')
phenodat<-read.csv('unblinded_pheno_sct_twin.csv',stringsAsFactors=FALSE)
#This file was created using 'geno_pheno_analysis_190215.Rmd' - i.e. main script for Newbury et al paper. In that script, IDs are blinded, but here we need to match up with CNV data, so retain the codes.

#need to strip out correct ID from the cnv files
#first of all remove _rpt from end of name
#also remove DB from start
#using cnvdatx as this seems more accurate, but ultimately should agree with cnvdat



for (i in 1:nrow(cnvdatx)){
  cnvdatx$id[i] <-  str_replace(cnvdatx$id[i],"_rpt", "")
  cnvdatx$id[i] <-  str_replace(cnvdatx$id[i],"DB", "")
    cnvdat$id[i] <-  str_replace(cnvdat$id[i],"_rpt", "")
  cnvdat$id[i] <-  str_replace(cnvdat$id[i],"DB", "")
}

#for SCTs, just retain first 3 digits
myrows<-which(cnvdatx$Group != 'twin')

cnvdatx$id[myrows]<-str_sub(cnvdatx$id[myrows],1,3)
cnvdat$id[myrows]<-str_sub(cnvdatx$id[myrows],1,3)
#NB these ids are characters, so need to alter phenodat,where id is numeric
#Need to add leading zeroes, so, e.g. 1 becomes 001

phenodat$id<-as.character(phenodat$record_id)
w<-which(as.numeric(phenodat$record_id)<10)
phenodat$id[w]<-paste0('0',phenodat$id[w])
w<-which(as.numeric(phenodat$record_id)<100)
phenodat$id[w]<-paste0('0',phenodat$id[w])

#order all files by id
genderdat <- genderdat[order(genderdat$record_id),] 
phenodat <- phenodat[order(phenodat$id),] 
cnvdat <- cnvdat[order(cnvdat$id),] 
cnvdatx <- cnvdatx[order(cnvdatx$id),] 
#align files and bolt together

#Just check we do have same cases in same rows
for(i in 1:nrow(cnvdat)){
  if(cnvdat$id[i] != cnvdatx$id[i]){
    print(cnvdat$id[i])
  }
  }



w<-which(phenodat$id %in% cnvdatx$id)

cnvpheno<-cbind(phenodat[w,2:44],cnvdat,cnvdatx)

cnvpheno$female<-0
w4<-which(cnvpheno$trisomy==1)
cnvpheno$female[w4]<-1

w3<-which(cnvpheno$Group=='twin')
w2<-which(genderdat$record_id %in% cnvpheno$id) #twins only!

cnvpheno$female[w3]<-genderdat$female[w2] #twins who are in cnvpheno assigned gender from genderdat

cnvpheno$groupfem<-paste0(cnvpheno$Group,cnvpheno$female)

#Do some renaming of cols
for (i in 56:64){
  colnames(cnvpheno)[i]<-paste0(colnames(cnvpheno)[i],'X')
}
colnames(cnvpheno)[63]<-'is_trisomy'
#Make an overall code for trisomy vs twin
cnvpheno$is_trisomy <-1
w<-which(cnvpheno$Group=='twin')
cnvpheno$is_trisomy[w]<-0 
```

We now have a file with the phenotype and genotype data together.  
First thing that is needed is Table 1, showing Ns in relation to ascertainment bias.
```{r table1}

cnvpheno$bias<-cnvpheno$famsplangconc #default from twin code for family bias

w<-which(cnvpheno$Group !='twin')
cnvpheno$bias[w]<-0 #default no bias for trisomies
w<-c(which(cnvpheno$why_tested==2),which(cnvpheno$why_tested==3))
cnvpheno$bias[w]<-1

mytab<- table(cnvpheno$groupfem,cnvpheno$bias)


```
The derivation of the global neurodevelopmental index is now provided in the paper in Table 2.
I have subtracted scores from 10, so that a high score represents good functioning. 

```{r globalneuro}

cnvpheno$gni_inv<-(10-cnvpheno$global_neurodev)

```

Just recreating Hayley's analysis of burden.

```{r burdentable}
w<-which(cnvpheno$bias==1)
biasset<-cnvpheno[w,]
nobiasset<-cnvpheno[-w,]

mytablea<-table(nobiasset$is_trisomy,nobiasset$total_cnvs)
print('Low bias groups: autosomes')
round(prop.table(mytablea,1)*100,1)
print('Low bias groups: Xchr')
mytablex<-table(nobiasset$is_trisomy,nobiasset$total_cnvsX)
round(prop.table(mytablex,1)*100,1)


#do pirate-plots
pirateplot(log(1+cnvpheno$total_cnvs)~cnvpheno$bias*cnvpheno$is_trisomy,data=cnvpheno)
pirateplot(log(1+cnvpheno$total_Kbp)~cnvpheno$bias*cnvpheno$is_trisomy,data=cnvpheno)

#As explained in the paper, the X chromosome is a bit of a nightmare if you have 3 Xs!
pirateplot(log(1+log(1+cnvpheno$total_KbpX))~cnvpheno$bias*cnvpheno$is_trisomy*cnvpheno$female,data=cnvpheno)
```
